### 2015-10-12 paper/main.tex
Snapshot + paper.

\documentclass{easychair}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage{amstext}
\usepackage{amsthm}
\usepackage{hyperref}

\usepackage{tabularx}
\usepackage{listings}
  \def\li{\lstinline}
  \lstset{
    language=caml,
    flexiblecolumns=false,
    showstringspaces=false,
    basicstyle=\ttfamily,
    framesep=5pt,
    numberstyle=\tiny,
    numbersep=5pt,
    escapeinside={(*}{*)},
    morekeywords={above,abstract,adopts,alias,and,as,assert,begin,below,builtin,consumes,data,do,downto,duplicable,dynamic,else,empty,end,exclusive,explain,fact,fail,flex,for,from,fun,give,if,in,let,match,mutable,open,pack,perm,preserving,rec,take,taking,then,to,type,unknown,val,value,while,with,witness},
    deletekeywords={value},
  }
\usepackage{minted}
  \newminted[ocaml]{ocaml}{mathescape}
\usepackage{mathpartir}
  % Improvements to Didier's mathpartir package: make all rule names and
  % references use textsc, and also make references to rules done using the \Rule
  % command clickable in a PDF.
  \let\TirName\textsc
  \renewcommand{\RefTirName}[1]{\hypertarget{#1}{\TirName {#1}}}
  \renewcommand{\DefTirName}[1]{\hyperlink{#1}{\TirName {#1}}}
  \let\DefRule\RefTirName
  \let\Rule\DefTirName

\newcommand{\fref}[1]{Figure~\ref{fig:#1}}
\newcommand{\R}{\ensuremath{\mathcal{R}}} % Rigid
\newcommand{\f}[1]{\ensuremath{\mathbf{#1}}} % Bold
\newcommand{\F}{\ensuremath{\mathcal{F}}} % Flexible
\newcommand{\V}{\ensuremath{\mathcal{V}}} % The prefix

\begin{document}

\title{Functional Pearl: the Proof Search Monad}
\titlerunning{The Proof Search Monad}

\author{Jonathan Protzenko}
\authorrunning{J. Protzenko}
\institute{
  Microsoft Research\\
  \email{protz@microsoft.com}}

\maketitle

\begin{abstract}
  We present the proof search monad, a set of combinators that allows one to
  write a proof search engine in a style that resembles closely the inference
  rules. The user calls functions such as \li+premise+, \li+prove+ or
  \li+choice+; the library then takes care of generating a derivation tree,
  which serves either as a witness (success) or as an error message (failure).
  Proof search engines written in this style enjoy: first, a one-to-one
  correspondence between the implementation and the theoretical rules, which
  makes manual inspection easier; second, proof witnesses ``for free'', which
  makes a verified-validation approach easier too.
\end{abstract}

\section{A minimal problem}

We consider conjunctions of equalities of the form
$\V x_1, \ldots, \V x_n.\ \bigwedge_k x_i = x_j$, that is, conjunctions of
equalities between variables. Some variables are \emph{flexible}, meaning they may be
substituted for other variables. Their binder $\V x_i$ is of the form $\F \f
x_i$ (and their are typeset in bold font). Other variables are \emph{rigid}, and
may not be substituted. Their binder $\V x_i$ is of the form $\R x_i$. The
problem consists in computing a substitution of \emph{flexible} variables that
evaluate the conjunction to \li+true+. If no such substitution exists, the
solver outputs nothing.

For instance, one may want to solve: $\R x, \R z, \F \f y.\ x = \f y \wedge z = z$.
A solution exists: the solver outputs $\sigma = \{ \f y \mapsto x \}$ as a valid
substitution that solves the input problem. However, if one attempts to solve:
$\R x, \R z, \F \f y.\ x = \f y \wedge \f y = z$, the solver fails to find a
proper substitution, and return nothing. Indeed, the first clause demands that
$\f y$ substitutes to $x$, meaning that the second clause becomes $x = z$, which
always evaluates to false ($x$ and $z$ are two distinct rigid variables).

In proof search, rigid variables stem from the right-elimination of universal
quantifiers, or left-elimination of existential quantifiers, which both result
in abstract variables. When left-eliminating a universal quantifier
(respectively right-eliminating an existential quantifier), one must provide an
argument to the type application (resp. an existential witness). If the argument
(resp. the witness) cannot be found on the spot, one typically uses
\emph{flexible variables}, which allow the proof search procedure to
\emph{defer} the choice until a later point in search (where the choice may be
guessed). Flexible variables are thus an \emph{implementation technique}.

In order to simplify the discussion, we skip elimination rules and quantifiers
altogether, and just assume that our input problem contains a combination of
flexible and rigid variables. Furthermore, we assume that any instantiation of a
flexible variable is legal. The only crucial point is that the solver outputs
something.

(The library that we describe here has been used in the implementation of the
Mezzo type-checker. There, all introduction and elimination rules are present;
furthermore, the implementation uses \emph{levels} to guarantee that only sound
instantiation choices are performed. This is, however, orthogonal to the present
discussion; the minimal problem we consider already conveys all the important
points, so we will skip quantifiers in the remainder of the discussion.)

Solving conjunctions of equalities requires three rules, which make up our
logic (\fref{proof-system}). \Rule{Refl} embodies the reflexivity axiom; \Rule{And}
highlights that the proof produces an \emph{output} $V'$ which is \emph{chained}
from one premise to another; \Rule{Inst} embodies the ``instantiation''
mechanism: in essence, if $y$ has not been instantiated so far, one
may instantiate it onto any other variable and perform the substitution in order
to prove the goal ($[\f y/x] P$ stands for ``replace $y$ with $x$ in $P$'').

\begin{figure}
  \centering
  \begin{mathpar}
    \inferrule[Refl]{
      \quad
    }{
      V \vdash x = x \dashv V
    }

    \inferrule[Inst]{
      x \in V_i \\
      V_1, \F(\f y = x), V_2 \vdash [\f y/x] P \dashv V'
    }{
      V_1, \F(\f y), V_2 \vdash P \dashv V'
    }

    \inferrule[And]{
      V \vdash P \dashv V' \\
      V' \vdash Q \dashv V''
    }{
      V \vdash P \wedge Q \dashv V''
    }
  \end{mathpar}
  \caption{Proof rules for the first system}
  \label{fig:proof-system}
\end{figure}

Once the proof tree has been laid out, one obtains an output (e.g. $V''$ in
\Rule{And}) where the instantiation choices (e.g. $\F (\f x_i = x_j)$) determine
the desired substitution (e.g. $\sigma = \{ \f x_i \mapsto x_j \}^*$).

\begin{figure}
  \centering
  \begin{ocaml}
type formula =                and descr =
| Equals of var * var           | Flexible
| And of formula * formula      | Rigid

and var = P.point             and env = descr P.state
  \end{ocaml}
  \caption{Formulas and environments}
  \label{fig:formulas}
\end{figure}

We implement proof search in OCaml (\fref{formulas}). The data type of formulas
is self-explanatory. Variables are implemented as equivalence classes in a
\emph{persistent} union-find data structure, which the module \li+P+ implements.
The $V$ parameter in our rules is embodied by the \li+env+ type
(``environment''), which is an input and an output to the solver. An
\li+env+ represents a given state of the union-find data structure, that is, a
set of equations between variables. This is another way of representing a
substitution, i.e. this is what we want our solver to compute.

The choice of a persistent union-find is irrelevant. All that matters is that we
pick a data structure that models substitutions, and that is \emph{persistent}.
Had we picked an explicit substitution instead of a union-find, the rest of the
discussion would have been the same.

\fref{solver} implements a solver for our minimal problem; written within the
\li+MOption+ monad, it returns either \li+Some env+ (in case a successful
substitution has been found), or \li+None+ if no solution exists. The solver is
complete.

\begin{figure}
  \centering
  \begin{ocaml}
module MOption = struct
  (* defines [return], [nothing] and [>>=] *)
end

let unify env v1 v2 =                         let rec solve env formula =
  match P.find v1 env, P.find v2 env with       match formula with
  | Flexible, Flexible                          | Equals (v1, v2) ->
  | Flexible, Rigid ->                              unify env v1 v2
      return (P.union v1 v2 env)                | And (f1, f2) ->
  | Rigid, Flexible ->                              solve env f1 >>= fun env ->
      return (P.union v2 v1 env)                    solve env f2
  | Rigid, Rigid ->
      if P.same v1 v2 env then
        return env
      else
        nothing
  \end{ocaml}
  \caption{Solver for the simplified problem}
  \label{fig:solver}
\end{figure}

The solver uses \li+MOption.>>=+ to sequence premises in the \li+And+ case. It
doesn't keep track of premises; it just ensures (thanks to \li+>>=+) that if the
first premise evaluates to \li+nothing+, the second premise is not evaluated,
since it is suspended behind a \li+fun+ expression (OCaml is a strict language).

\section{Building derivations}

There are two shortcomings with this solver. First, the \li+unify+ sub-routine
conflates several rules of the logic together. Indeed, the \li+return (P.union ...)+
expression hides a combination of \Rule{Inst} and \Rule{Refl}. Second, we
have no way to replay the proof to verify it independently. One may argue that
in this simplified example, one can just apply the substitution to the original
formula and verify that all the clauses are of the form $x = x$, without the
need for a proof tree. In the general case, however, the proof tree contains the
elimination witnesses for quantifiers; this allows one to independently verify a
proof without relying on inference techniques, such as flexible variables.

\subsection{Defining proof trees}

One way to make the solver better is to make sure each step it performs
corresponds in an obvious manner to the application of an admissible rule. To
that effect, we define the data type of all three rules in our system, which we
apply to the functor of \emph{proof trees} (\fref{proof-trees}).

\begin{figure}
  \centering
\begin{ocaml}
module type LOGIC = sig     module MakeProofTree (L: LOGIC) = struct
  type formula                type derivation = L.formula * rule
  type rule_name              and rule = L.rule_name * premises
end                           and premises = Premises of derivation list
                            end

module MyLogic = struct
  type formula = ... (* as before *)
  type rule_name = R_And | R_Refl | R_Inst
end
module MyProofTree = MakeProofTree(MyLogic)
\end{ocaml}
  \caption{The functor of proof trees}
  \label{fig:proof-trees}
\end{figure}

A \li+derivation+ tree is a pair of a \li+formula+ (the goal we wish the prove)
and a \li+rule+ (that we apply in order to prove the goal). A \li+rule+ has a name
and \li+premises+; the \li+premises+ type is simply a \li+derivation list+ (the
\li+Premises+ constructor is here to prevent a non-constructive type
abbreviation). When using the library, the client is expected to make sure
that each \li+rule_name+ is paired with the proper
number of premises (0 for \Rule{Refl}, 1 for
\Rule{Inst} and 2 for \Rule{And}); this is not enforced by the type system.

In the (simplified) sketch from \fref{proof-trees}, rule names are just constant
constructors, since the rule parameters (such as $x$ and $\f y$ in \Rule{Inst})
can be recovered from the \li+formula+. In the general case, the various
constructors of \li+rule_name+ do have parameters that record how one specific
rule was instantiated.

\subsection{Proof tree combinators}

We previously used the \li+>>=+ operator from the \li+MOption+ monad in order to
chain premises (\fref{solver}). We now need a new operator, that not only
\emph{binds} the result (i.e. stops evaluating premises after a failure, as
before), but also \emph{records} the premises in sequence, in order to build a
proper derivation. The former is still faithfully implemented by the option
monad; the latter is implemented by the writer monad.

Computations in the writer monad return a result (of type \li+'a+) along with a log of
elements (of type \li+L.a+). The (usual) \li+>>=+ and \li+return+ combinators operate on
the result part of the computation, while the (new) \li+tell+ combinator
operates on the logging part of the computation. This \li+tell+ combinator
appends a new element to the log. Appending elements to the log is done by way of the \li+MONOID+
module type, which essentially demands a value for the empty log, and a function to
append new entries into the log.

In order to get a new \li+>>=+ operator that combines the features of
the option and writer monads, we apply the \li+WriterT+ monad transformer to the
\li+MOption+ monad (\fref{writer}) and obtain \li+MWriter+, a monad whose
computations represent a sequence of derivations (the premises we have proved so far)
along with a result (the \li+env+ that we chain through the premises). These
computations are wrapped in \li+MOption.m+, that is, are wrapped within an
\li+option+ to account for a possible proof failure.

\begin{figure}
  \centering
\begin{ocaml}
module WriterT (M: MONAD) (L: MONOID): sig      module L = struct
  type 'a m = (L.a * 'a) M.m                      type a = MyProofTree.derivation list
  val return: 'a -> 'a m                          let empty = []
  val ( >>= ): 'a m -> ('a -> 'b m) -> 'b m       let append = List.append
                                                end
  val tell: L.a -> unit m
end = ...

module MWriter = WriterT(MOption)(L)
\end{ocaml}
  \caption{The writer monad transformer}
  \label{fig:writer}
\end{figure}

\bibliographystyle{abbrvnat}
\bibliography{local}

\end{document}